Interisland movements of woodland caribou in Newfoundland

### 2020-02-21 paper/manuscript.Rmd
more pred stuff

## Predator avoidance
Predator avoidance could also contribute to swimming behaviour of caribou
(Miller 2002). During calving, female caribou are known to swim to small islands
in freshwater lakes give birth, or with their calves after birth, to avoid
predation [@Bergerud_1987]. The use of shoreline habitat in summer is considered
an effective anti-predator calving strategy by allowing for swimming to islands
or peninsulas as a predation avoidance flight response [@Bergerud_1990]. The use
of water, and by association occasionally swimming, is therefore part of the
fine-scale interactions between caribou and their predators and the use of
islands appears to be an effective anti-predator strategy.

Historic predators of caribou such as wolves (*Canis lupus*) and black bears
(*Ursus americanus*) are not present in the Fogo Island archipelago; coyotes
(*Canis latrans*) are the top predator in this region. Predation by coyotes on
adult caribou is relatively rare (ref?), so the potential predator effect on
caribou swimming behaviour is likely to be a flight response to non-lethal
encounters. Calves are susceptible to coyotes, particularly young calves, but
the energetic costs to oceanic swimming are likely to be far greater for small
calves.

@@

## Individual variation 
Three of 29 GPS-collared female caribou swam between islands in the Fogo
archipelago. These swimming events were relatively frequent, on average occuring
once every 24 days within our date range. However, we identified zero swimming
events for the other 26 collared females. It remains unclear why this behaviour
is seemingly common among 10% of our marked population, while 90% never engaged
in it. Our initial direct observation of swimming was an adult male, but as we
only have GPS data for females, we are unable to make conclusions on the
prevalence of this behaviour among male caribou. It is theoretically possible
that these other females moved rapidly back and forth between islands within the
2 hour fix rate of our GPS collars, which we would be unable to detect. We
consider this highly unlikely given the range of residency times for caribou on
any given island (min - max: XX - XX).
<!-- What is the typical residency time? Is this a true statement or am I just making this up? -->

>>>>

## Predator avoidance
Predator avoidance could also contribute to swimming behaviour of caribou
(Miller 2002) <!-- find doi for this article -->. During calving, female caribou
are known to swim to small islands in freshwater lakes give birth, or with their
calves after birth, to avoid predation [@Bergerud_1987]. The use of shoreline
habitat in summer is considered an effective anti-predator calving strategy by
allowing for swimming to islands or peninsulas as a predation avoidance flight
response [@Bergerud_1990]. The use of water, and by association occasionally
swimming, is therefore part of the fine-scale interactions between caribou and
their predators and the use of islands appears to be an effective anti-predator
strategy.

Historic predators of caribou such as wolves (*Canis lupus*) and black bears
(*Ursus americanus*) are not present in the Fogo Island archipelago; coyotes
(*Canis latrans*) are the top predator in this region. Predation by coyotes on
adult caribou is relatively rare (ref?), so the potential effect of predators on
caribou swimming behaviour is likely to be a flight response to non-lethal
encounters. Calves are susceptible to coyotes, particularly young calves, but
the energetic costs to oceanic swimming for calves is likely to be far greater for small
calves compared to adults.

@@

## Individual variation 
In total, 3 of 29 collared female caribou swam between islands in the Fogo
archipelago. These swimming events were relatively frequent, on average occuring
once every 24 <!--double check--> days during ice-free seasons. However, we
identified zero swimming events for the other 26 collared females. It remains
unclear why this behaviour is seemingly common among 10% of our marked
population, while 90% never engaged in it. Our initial direct observation of
swimming was an adult male, but as we only have GPS data for females, we are
unable to make conclusions on the prevalence of this behaviour among male
caribou. It is theoretically possible that these other females moved rapidly
back and forth between islands within the 2 hour fix rate of our GPS collars,
which we would be unable to detect. We consider this highly unlikely given the
range of residency times for caribou on any given island (min - max: XX - XX).
<!-- What is the typical residency time? Is this a true statement or am I just making this up? -->
